Incentivizing Exploration with Unbiased History

### 2018-11-01 llevel.tex
T_G -> alex's macro
q_a -> \fdpN
fun fact: we now have at most 30 q's in the paper

For $a \in \{1,2\}$, define $q_a$ to be the expected number of arm $a$ pulls in one run of \ALGG used in the first level. By Lemma \ref{lem:greedy}, we know $\GdP \leq q_a \leq \GdT$ For group $G_{1,u,v}$, define $W_1^{a,u,v}$ to be the event that the number of arm $a$ pulls in this group is between $q_a T_1- \GdT \sqrt{T_1\log(T)}$ and $q_a T_1 + \GdT \sqrt{T_1\log(T)}$. By Chernoff bound,
\[
\Pr[W_1^{a,u,v}] \geq 1-2\exp(-2\log(T)) \geq 1-2/T^2.
\]
Define $W_1$ to be the intersection of all these events (i.e. $W_1 = \bigcap_{a,u,v}W_1^{a,u,v}$). By union bound, we have
\[
\Pr[W_1] \geq 1- \frac{4S^2}{T^2}.
\]

@@

For first-level groups $G_{1,u,1},...,G_{1,u,S}$ and arm $a$, imagine there is a tape of enough arm $a$ pulls sampled before the recommendation policy starts and these samples are revealed one by one whenever agents in these groups pull arm $a$. Define $W_4^{u,a,high}$  to be the event that first $q_a T_1 S$ pulls of arm $a$ in the tape has empirical mean at least $\mu_a + 1/\sqrt{q_a T_1 S}$ and define $W_4^{u,a,low}$  to be the event that first $q_a T_1S$ pulls of arm $a$ in the tape has empirical mean at most $\mu_a - 1/\sqrt{q_a T_1S }$. By Berry-Esseen Theorem and $\mu_a \in [1/3,2/3]$, we have
\[
\Pr[W_4^{u,a,high}] \geq (1-\Phi(1/2)) - \frac{5}{\sqrt{q_aT_1S}} > 1/4.
\]
The last inequality follows when $T$ is larger than some constant.
Similarly we also have 
\[
\Pr[W_4^{u,a,low}] > 1/4.
\]
Since $W_4^{u,a,high}$ is independent with $W_4^{u,3-a,low}$, we have
\[
\Pr[W_4^{u,a,high} \cap W_4^{u,3-a,low}] =\Pr[W_4^{u,a,high}] \cdot  \Pr[W_4^{u,3-a,low}]>(1/4)^2 = 1/16.
\]
Now define $W^{a}_4$ as $\bigcup_u (W_4^{u,a,high} \cap W_4^{u,3-a,low})$. Notice that $(W_4^{u,a,high} \cap W_4^{u,3-a,low})$ are independent across different $u$'s. So we have
\[
\Pr[W^{a}_4] \geq 1- (1-1/16)^S \geq 1 -1/T^2.
\]
Finally we define $W_4$ as $\bigcap_{a} W^{a}_4$. By union bound,
\[
\Pr[W_4] \geq 1- 2/T^2.
\]
\end{itemize}

@@

By event $W_1$, we know that in each first-level group, there are at least $q_a T_1- \GdT \sqrt{T_1\log(T)}$ pulls of arm $a$. We prove in the next claim that there are enough pulls of both arms in higher levels if $\mu_1-\mu_2$ is small enough. For notation convenience, we set $\varepsilon_0 = 1$, $\varepsilon_1 = \frac{1}{4\sqrt{q_aT_1S}} + \frac{1}{4\sqrt{q_{3-a} T_1S}}$ and $\varepsilon_l = 1/(4\sqrt{T_lS})$ for $l \geq 2$.

@@

We prove the above via induction on $l$. %For notation convenience, define $\bar{\mu}^{l,u}_a$  to be the empirical mean of arm $a$ pulls in the history observed by agents in groups $G_{l,1,u},...,G_{l,S,u}$ and $\Gamma_{l,1,u},...,\Gamma_{l,S,u}$ (they observe the same history).
We start by the base case when $l=2$. For each arm $a$, $W_4$ implies there exists $u_a$ such that $W^{u_a,a,high}_4$ and $W^{u_a,3-a,low}_4$ happen. For an agent $t$  in groups $G_{2,1,u_a},...,G_{2,S,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,S,u_a}$.
$W_4^{u_a,a,high}$,  $W_1^{a,u_a,v}$ and $W_2$ together imply that 
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(q_aT_1S \cdot \frac{1}{\sqrt{q_aT_1S}} - \GdT \sqrt{T_1\log(T)} S\cdot \sqrt{\frac{3\log(T)}{ \GdT \sqrt{T_1\log(T)}S}} \right) \cdot \frac{1}{(q_a T_1+ \GdT \sqrt{T_1\log(T)})S} \\
&> \mu_a + \frac{1}{4\sqrt{q_aT_1S}}.
\end{align*}
The second last inequality holds when $T$ is larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t< \mu_{3-a}   - \frac{1}{4\sqrt{q_{3-a} T_1S}}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a} + \frac{1}{4\sqrt{q_aT_1S}} + \frac{1}{4\sqrt{q_{3-a} T_1S}}\\
&\geq -\varepsilon_1+ \frac{1}{4\sqrt{q_aT_1S}} + \frac{1}{4\sqrt{q_{3-a} T_1S}}\\
&\geq \frac{1}{8\sqrt{q_aT_1S}} + \frac{1}{8\sqrt{q_{3-a} T_1S}}.
\end{align*}
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{c_m}{\sqrt{q_aT_1S/2}} - \frac{c_m}{\sqrt{q_{3-a} T_1S/2}}\\
&> \frac{1}{8\sqrt{q_aT_1S}} + \frac{1}{8\sqrt{q_{3-a} T_1S}} -   \frac{c_m}{\sqrt{q_aT_1S/2}} - \frac{c_m}{\sqrt{q_{3-a} T_1S/2}}\\
&>0.
\end{align*}
The last inequality holds since $c_m$ is a small enough constant defined in Assumption \ref{ass:embehave}. Therefore we know agents in groups $G_{2,1,u_a},...,G_{2,S,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,S,u_a}$ all pull arm $a$.

@@

\begin{itemize}
\item $\varepsilon_{l-1} S \leq \mu_1 - \mu_2 \leq \varepsilon_{l-2}$ for $l = 2$(i.e. $\varepsilon_1S\leq \mu_1 - \mu_2 \leq \varepsilon_0$): We know agents in level at least 2 will observe at least $q_aT_1/2$ pulls of arm $a$ for $a \in \{1,2\}$. By $W_2$, for any agent in level at least 2, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{Sq_aT_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{c_m}{\sqrt{Sq_1T_1/2}} -  \frac{c_m}{\sqrt{Sq_2T_1/2}}\\
&\geq \mu_1 -\mu_2-\sqrt{\frac{3\log(T)}{Sq_1T_1/2}} - \sqrt{\frac{3\log(T)}{Sq_2T_1/2}}- \frac{c_m}{\sqrt{Sq_1T_1/2}} -  \frac{c_m}{\sqrt{Sq_2T_1/2}}\\
&\geq\frac{\sqrt{S}}{4\sqrt{q_1T_1}} +  \frac{\sqrt{S}}{4\sqrt{q_2T_1}}-\sqrt{\frac{3\log(T)}{Sq_1T_1/2}} - \sqrt{\frac{3\log(T)}{Sq_2T_1/2}}- \frac{c_m}{\sqrt{Sq_1T_1/2}} -  \frac{c_m}{\sqrt{Sq_2T_1/2}}\\
&>0.
\end{align*}
Therefore agents in level at least 2 will all pull arm 1.

@@

\item $\varepsilon_{l-2} < \mu_1 - \mu_2 < \varepsilon_{l-2}S$ for $l =3$ (i.e. $\varepsilon_1 < \mu_1 - \mu_2 < \varepsilon_1S$): By Claim \ref{clm:l2_explore}, for any agent $t$ in level at least $3$, that agent will observe at least $T_1q_aS^2/2$ arm $a$ pulls (just from the first level). By $W_2$, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{S^2q_aT_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{c_m}{\sqrt{S^2q_1T_1/2}}-\frac{c_m}{\sqrt{S^2q_2T_1/2}}  \\
&\geq \mu_1 -\mu_2 -  \sqrt{\frac{3\log(T)}{S^2q_1T_1/2}}- \sqrt{\frac{3\log(T)}{S^2q_2T_1/2}}- \frac{c_m}{\sqrt{S^2q_1T_1/2}}-\frac{c_m}{\sqrt{S^2q_2T_1/2}}  \\
&\geq\frac{1}{4\sqrt{Sq_1T_1}} + \frac{1}{4\sqrt{Sq_2T_1}}  -  \sqrt{\frac{3\log(T)}{S^2q_1T_1/2}}- \sqrt{\frac{3\log(T)}{S^2q_2T_1/2}}- \frac{c_m}{\sqrt{S^2q_1T_1/2}}-\frac{c_m}{\sqrt{S^2q_2T_1/2}}  \\
&>0.
\end{align*}
Therefore agents in level at least 3 will all pull arm 1.

@@

By Claim \ref{clm:l2_exploit}, we know that the expected regret conditioned the intersection of clean events is at most 
\begin{align*}
&\max\left( T_1T_GS^2 , \max_{l \geq 2} \varepsilon_{l-1}S(T_1T_GS^2 + T_2 S^3 + \cdots + T_lS^3)\right) \\
\leq & \max\left( T_1T_GS^2 , \max_{l \geq 2} 2 \varepsilon_{l-1} T_l S^4 \right)\\
= &O\left(T^{2^{L-1}/(2^L-1)} \log^2(T) \right).
\end{align*}
\end{proof}

>>>>

For $a \in \{1,2\}$, define $\fdpN$ to be the expected number of arm $a$ pulls in one run of \ALGG used in the first level. By Lemma \ref{lem:greedy}, we know $\GdP \leq \fdpN \leq \GdT$ For group $G_{1,u,v}$, define $W_1^{a,u,v}$ to be the event that the number of arm $a$ pulls in this group is between $\fdpN T_1- \GdT \sqrt{T_1\log(T)}$ and $\fdpN T_1 + \GdT \sqrt{T_1\log(T)}$. By Chernoff bound,
\[
\Pr[W_1^{a,u,v}] \geq 1-2\exp(-2\log(T)) \geq 1-2/T^2.
\]
Define $W_1$ to be the intersection of all these events (i.e. $W_1 = \bigcap_{a,u,v}W_1^{a,u,v}$). By union bound, we have
\[
\Pr[W_1] \geq 1- \frac{4S^2}{T^2}.
\]

@@

For first-level groups $G_{1,u,1},...,G_{1,u,S}$ and arm $a$, imagine there is a tape of enough arm $a$ pulls sampled before the recommendation policy starts and these samples are revealed one by one whenever agents in these groups pull arm $a$. Define $W_4^{u,a,high}$  to be the event that first $\fdpN T_1 S$ pulls of arm $a$ in the tape has empirical mean at least $\mu_a + 1/\sqrt{\fdpN T_1 S}$ and define $W_4^{u,a,low}$  to be the event that first $\fdpN T_1S$ pulls of arm $a$ in the tape has empirical mean at most $\mu_a - 1/\sqrt{\fdpN T_1S }$. By Berry-Esseen Theorem and $\mu_a \in [1/3,2/3]$, we have
\[
\Pr[W_4^{u,a,high}] \geq (1-\Phi(1/2)) - \frac{5}{\sqrt{\fdpN T_1S}} > 1/4.
\]
The last inequality follows when $T$ is larger than some constant.
Similarly we also have 
\[
\Pr[W_4^{u,a,low}] > 1/4.
\]
Since $W_4^{u,a,high}$ is independent with $W_4^{u,3-a,low}$, we have
\[
\Pr[W_4^{u,a,high} \cap W_4^{u,3-a,low}] =\Pr[W_4^{u,a,high}] \cdot  \Pr[W_4^{u,3-a,low}]>(1/4)^2 = 1/16.
\]
Now define $W^{a}_4$ as $\bigcup_u (W_4^{u,a,high} \cap W_4^{u,3-a,low})$. Notice that $(W_4^{u,a,high} \cap W_4^{u,3-a,low})$ are independent across different $u$'s. So we have
\[
\Pr[W^{a}_4] \geq 1- (1-1/16)^S \geq 1 -1/T^2.
\]
Finally we define $W_4$ as $\bigcap_{a} W^{a}_4$. By union bound,
\[
\Pr[W_4] \geq 1- 2/T^2.
\]
\end{itemize}

@@

By event $W_1$, we know that in each first-level group, there are at least $\fdpN T_1- \GdT \sqrt{T_1\log(T)}$ pulls of arm $a$. We prove in the next claim that there are enough pulls of both arms in higher levels if $\mu_1-\mu_2$ is small enough. For notation convenience, we set $\varepsilon_0 = 1$, $\varepsilon_1 = \frac{1}{4\sqrt{\fdpN T_1S}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1S}}$ and $\varepsilon_l = 1/(4\sqrt{T_lS})$ for $l \geq 2$.

@@

We prove the above via induction on $l$. %For notation convenience, define $\bar{\mu}^{l,u}_a$  to be the empirical mean of arm $a$ pulls in the history observed by agents in groups $G_{l,1,u},...,G_{l,S,u}$ and $\Gamma_{l,1,u},...,\Gamma_{l,S,u}$ (they observe the same history).
We start by the base case when $l=2$. For each arm $a$, $W_4$ implies there exists $u_a$ such that $W^{u_a,a,high}_4$ and $W^{u_a,3-a,low}_4$ happen. For an agent $t$  in groups $G_{2,1,u_a},...,G_{2,S,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,S,u_a}$.
$W_4^{u_a,a,high}$,  $W_1^{a,u_a,v}$ and $W_2$ together imply that 
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(\fdpN T_1S \cdot \frac{1}{\sqrt{\fdpN T_1S}} - \GdT \sqrt{T_1\log(T)} S\cdot \sqrt{\frac{3\log(T)}{ \GdT \sqrt{T_1\log(T)}S}} \right) \cdot \frac{1}{(\fdpN T_1+ \GdT \sqrt{T_1\log(T)})S} \\
&> \mu_a + \frac{1}{4\sqrt{\fdpN T_1S}}.
\end{align*}
The second last inequality holds when $T$ is larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t< \mu_{3-a}   - \frac{1}{4\sqrt{\fdpN[3-a] T_1S}}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a} + \frac{1}{4\sqrt{\fdpN T_1S}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1S}}\\
&\geq -\varepsilon_1+ \frac{1}{4\sqrt{\fdpN T_1S}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1S}}\\
&\geq \frac{1}{8\sqrt{\fdpN T_1S}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1S}}.
\end{align*}
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{c_m}{\sqrt{\fdpN T_1S/2}} - \frac{c_m}{\sqrt{\fdpN[3-a] T_1S/2}}\\
&> \frac{1}{8\sqrt{\fdpN T_1S}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1S}} -   \frac{c_m}{\sqrt{\fdpN T_1S/2}} - \frac{c_m}{\sqrt{\fdpN[3-a] T_1S/2}}\\
&>0.
\end{align*}
The last inequality holds since $c_m$ is a small enough constant defined in Assumption \ref{ass:embehave}. Therefore we know agents in groups $G_{2,1,u_a},...,G_{2,S,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,S,u_a}$ all pull arm $a$.

@@

\begin{itemize}
\item $\varepsilon_{l-1} S \leq \mu_1 - \mu_2 \leq \varepsilon_{l-2}$ for $l = 2$(i.e. $\varepsilon_1S\leq \mu_1 - \mu_2 \leq \varepsilon_0$): We know agents in level at least 2 will observe at least $\fdpN T_1/2$ pulls of arm $a$ for $a \in \{1,2\}$. By $W_2$, for any agent in level at least 2, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{S\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{c_m}{\sqrt{S\fdpN[1]T_1/2}} -  \frac{c_m}{\sqrt{S\fdpN[2]T_1/2}}\\
&\geq \mu_1 -\mu_2-\sqrt{\frac{3\log(T)}{S\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{S\fdpN[2]T_1/2}}- \frac{c_m}{\sqrt{S\fdpN[1]T_1/2}} -  \frac{c_m}{\sqrt{S\fdpN[2]T_1/2}}\\
&\geq\frac{\sqrt{S}}{4\sqrt{\fdpN[1]T_1}} +  \frac{\sqrt{S}}{4\sqrt{\fdpN[2]T_1}}-\sqrt{\frac{3\log(T)}{S\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{S\fdpN[2]T_1/2}}- \frac{c_m}{\sqrt{S\fdpN[1]T_1/2}} -  \frac{c_m}{\sqrt{S\fdpN[2]T_1/2}}\\
&>0.
\end{align*}
Therefore agents in level at least 2 will all pull arm 1.

@@

\item $\varepsilon_{l-2} < \mu_1 - \mu_2 < \varepsilon_{l-2}S$ for $l =3$ (i.e. $\varepsilon_1 < \mu_1 - \mu_2 < \varepsilon_1S$): By Claim \ref{clm:l2_explore}, for any agent $t$ in level at least $3$, that agent will observe at least $T_1\fdpN S^2/2$ arm $a$ pulls (just from the first level). By $W_2$, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{S^2\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{c_m}{\sqrt{S^2\fdpN[1]T_1/2}}-\frac{c_m}{\sqrt{S^2\fdpN[2]T_1/2}}  \\
&\geq \mu_1 -\mu_2 -  \sqrt{\frac{3\log(T)}{S^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{S^2\fdpN[2]T_1/2}}- \frac{c_m}{\sqrt{S^2\fdpN[1]T_1/2}}-\frac{c_m}{\sqrt{S^2\fdpN[2]T_1/2}}  \\
&\geq\frac{1}{4\sqrt{S\fdpN[1]T_1}} + \frac{1}{4\sqrt{S\fdpN[2]T_1}}  -  \sqrt{\frac{3\log(T)}{S^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{S^2\fdpN[2]T_1/2}}- \frac{c_m}{\sqrt{S^2\fdpN[1]T_1/2}}-\frac{c_m}{\sqrt{S^2\fdpN[2]T_1/2}}  \\
&>0.
\end{align*}
Therefore agents in level at least 3 will all pull arm 1.

@@

By Claim \ref{clm:l2_exploit}, we know that the expected regret conditioned the intersection of clean events is at most 
\begin{align*}
&\max\left( T_1\GdT S^2 , \max_{l \geq 2} \varepsilon_{l-1}S(T_1\GdT S^2 + T_2 S^3 + \cdots + T_lS^3)\right) \\
\leq & \max\left( T_1\GdT S^2 , \max_{l \geq 2} 2 \varepsilon_{l-1} T_l S^4 \right)\\
= &O\left(T^{2^{L-1}/(2^L-1)} \log^2(T) \right).
\end{align*}
\end{proof}
